Locality and Computational Reliability of Linear Response Calculations for Molecular Systems

### 2018-07-12 AnalyticStructure_main.tex
Update on Overleaf.

In Fig.~\ref{CO_exc} and \ref{C6H6_exc} we plot the DOS of the Liouvillian for the $CO$  and Benzene molecules, 
for two different numerical setups of increasing size of the simulation domain. 
The figures show clearly that all the excitations of energy $\Omega_a$ lying below the 
molecule ionization potential (IP) have a clear localized behavior, as their energy does not vary with the size of the simulation domain.
On the other hand, above the first ionization threshold the excitation energies have an energy that strongly depends on the computational
setup, which is in line with the expected pseudo-continuum character of this sector. 
In addition, for the benzene molecule it is also possible to 
identify excitations of energy above IP that still possess a localized behavior, fulfilling the condition $\Omega_a < |\eps_a|$.
In Fig.~\ref{c6h6_excLand} we collect the excitations of the benzene molecule following this criterion:
the agreement with the expected analytic structure presented in Fig.~\ref{ExcitationLandscape} is remarkable.

@@

We may then decompose the linear response functional of Eq.~\eqref{LinearResponseFunctDef1}
for any operator $\op O$ as $\braket{\delta\op O}_\Phi=\braket{\delta\op O}^D_\Phi +
\braket{\delta\op O}^C_\Phi$, where
\be
\braket{\delta\op O}^{D,C}_\Phi(\omega) = 
2 \mathrm{Re}\left(\bra{\psi_p} \op O \fscd{D,C} \right)\;.
\ee
For instance, the static polarizability tensor $\alpha_{ij}$ can be easily expressed
from the above quantities as
\be
\alpha_{ij} = \sint \dd a \frac{w^a_{ij}}{\Omega_a^2}\,, \quad w^a_{ij} \equiv \sum_{p\in a} \varphi_p^a[\op r_i]\varphi_p^a[\op r_j]\;.
\ee
Such value can be straightforwardly compared with results of Eq.~\eqref{staticalpha}, where the localized FS is obtained from GS techniques as described in Sec.~\ref{FluctuationState}.
This is the well-know $S_{-2}$ sum-rule\cite{TDDFTBOOK}, that express the value of the static polarizability
in terms of the molecule's oscillator strengths.

@@

\begin{figure}
\includegraphics[scale=0.6]{CO_statPolvsExc.eps}
\caption{\label{co_AlphaExc}(Color online) Convergence of the statical polarizability of the $CO$ molecule w.r.t the number of excitations.}
\end{figure}
Hence, contrary to GS, \emph{for LR calculation,
even when the response density is localized, the contribution coming from the 
continuum sector of the excitations can never be neglected}. 
In other terms, we \emph{cannot} exclude that a fluctuation state, even if genuinely localized, will have a nonzero projection
in the essential part of the Liouvillian spectrum.
This fact shows that it is \emph{impossible} to restrict the spectral representation
of $\chi(\omega)$ only localized, discrete, \emph{code-independent} excitations, even for representing LR results in the $\omega=0$ limit.

>>>>

In Fig.~\ref{CO_exc} and \ref{C6H6_exc} we plot the DOS of the Liouvillian for the $CO$  and Benzene molecules, 
for two different numerical setups of increasing size of the simulation domain. 
The figures show clearly that all the excitations of energy $\Omega_a$ lying below the 
molecule ionization potential (IP) have a localized behavior, as their energy does not vary with the size of the simulation domain.
On the other hand, above the first ionization threshold the excitations have an energy that strongly depends on the computational
setup, which is in line with the expected pseudo-continuum character of this sector. 
In addition, for the benzene molecule it is also possible to 
identify excitations of energy above IP that still possess a localized behavior, fulfilling the condition $\Omega_a < |\eps_a|$.
In Fig.~\ref{c6h6_excLand} we collect the excitations of the benzene molecule following this criterion:
the agreement with the expected analytic structure presented in Fig.~\ref{ExcitationLandscape} is remarkable.

@@

We may then decompose the linear response functional of Eq.~\eqref{LinearResponseFunctDef1}
for any operator $\op O$ as $\braket{\delta\op O}_\Phi=\braket{\delta\op O}^D_\Phi +
\braket{\delta\op O}^C_\Phi$, where
\be
\braket{\delta\op O}^{D,C}_\Phi(\omega) = 
2 \mathrm{Re}\left(\bra{\psi_p} \op O \fscd{D,C} \right)\;.
\ee
For instance, the static polarizability tensor $\alpha_{ij}$ can be easily expressed
from the above quantities as
\be\lb{sminus2}
\alpha_{ij} = \sint \dd a \frac{w^a_{ij}}{\Omega_a^2}\,, \quad w^a_{ij} \equiv \sum_{p\in a} \varphi_p^a[\op r_i]\varphi_p^a[\op r_j]\;.
\ee
Such value can be straightforwardly compared with results of Eq.~\eqref{staticalpha}, where the localized FS is obtained from GS techniques as described in Sec.~\ref{FluctuationState}.
This is the well-know $S_{-2}$ sum-rule\cite{TDDFTBOOK}, that express the value of the static polarizability
in terms of the molecule's oscillator strengths.

@@

\begin{figure}
\includegraphics[scale=0.6]{CO_statPolvsExc.eps}
\caption{\label{co_AlphaExc}(Color online) Convergence of the statical polarizability $\alpha_{zz}$ of the $CO$ molecule, in a box of $11.5$ \AA w.r.t to excitations considered in the integral of Eq.~\eqref{sminus2}. The reference value obtained from Eq.~\eqref{staticalpha} is represented as a horizontal red dashed line. The value of the IP energy is indicated by a vertical black dashed line. For each excitation energy, the values of the oscillator strengths $w^a_{zz}$ of Eq.~\eqref{sminus2} are also plotted. }
\end{figure}
Hence, contrary to GS, \emph{for LR calculation,
even when the response density is localized, the contribution coming from the 
continuum sector of the excitations can never be neglected}. 
In other terms, we \emph{cannot} exclude that a fluctuation state, even if genuinely localized, will have a nonzero projection
in the essential part of the Liouvillian spectrum.
This fact shows that it is \emph{impossible} to restrict the spectral representation
of $\chi(\omega)$ only localized, discrete, \emph{code-independent} excitations, even for representing LR results in the $\omega=0$ limit.
